---
title: "Regional Diversity"
author: "vignettes/regional_diversity.Rmd"
date: "18/8/2022"
output: html_document
---


```{r}
library(obisindicators)
library(dggridR)  # remotes::install_github("r-barnes/dggridR")
library(dplyr)
library(sf)
library(tidyverse)
```


# Get regional biological occurrences
We will use de full OBIS dataset available at https://obis.org/data/access and filter data within the extension desired.

```{r}
occ <- 
  
  #get world data
  occ_1M %>% 
  
  #filter latitude
  filter(between(decimalLatitude, -1.3, 1.7)) %>% 
  
  #filter longitude
  filter(between(decimalLongitude, -92, -89.5))  

```

# Create a discrete global grid
Create an ISEA discrete global grid of resolution 15 using the dggridR package:

```{r}
dggs <- dgconstruct(projection = "ISEA", topology = "HEXAGON", res = 11)
#the higher the res, smaller the hexagons
```

Then assign cell numbers to the occurrence data:

```{r}
occ$cell <- dgGEO_to_SEQNUM(dggs, occ$decimalLongitude, occ$decimalLatitude)[["seqnum"]]
```

# Calculate indicators
The following function calculates the number of records, species richness, Simpson index, Shannon index, Hurlbert index (n = 50), and Hill numbers for each cell.

Perform the calculation on species level data:

```{r}
idx <- calc_indicators(occ)
```

Add cell geometries to the indicators table (idx):

```{r}
grid <- dgcellstogrid(dggs, idx$cell) %>% 
  st_wrap_dateline() %>%
  rename(cell = seqnum) %>%
  left_join(
    idx,
    by = "cell")
```


# Plot maps of indicators

We will personalize the gmap_indicator function.

```{r}
gmap_indicator <- function(
    grid, column = "shannon", label = "Shannon index", trans = "identity",
    crs="+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"){

  world <- rnaturalearth::ne_countries(scale = "large",type = "countries", returnclass = "sf")
  bb <- sf::st_bbox(
    st_transform(grid, crs))


  ggplot() +
    geom_sf(
      data = grid, aes_string(
        fill = column, geometry = "geometry"), lwd = 0) +
    viridis::scale_color_viridis(
      option = "viridis", na.value = "white",
      limits = c(0,5),
      name = label, trans = trans) +
    viridis::scale_fill_viridis(
      option = "viridis", na.value = "white",
      limits = c(0,5),
      name = label, trans = trans, alpha=0.75) +
    geom_sf(
      data = world, fill = "grey30", 
      color = "white",lwd=0.5) +
    theme_bw() +
    xlab("") + ylab("") +
    coord_sf(
      crs  = crs,
      xlim = bb[c("xmin","xmax")],
      ylim = bb[c("ymin","ymax")])
}
```


Resulting indicator in map form.

```{r shannon}
# Shannon index
gmap_indicator(grid, "shannon", label = "Shannon index")
```